Molecular transport in human pial perivasculature

### 2024-11-14 paper/supplementary_main.tex
Complete section A.1.2

Also in~\eqref{eq:3d_pde}, the notation $\overline{c}$ denotes a lateral average of the concentration over the outer perivascular surfaces, defined for each centerline $\Lambda^i$ and each point $s \in \Lambda^i$ by
\begin{equation}
  \overline{c}(s) = \frac{1}{P(s)} \int_{\partial \Theta_2^i(s)} c 
\end{equation}
where $\partial \Theta_2^i(s)$ is the outer boundary of the cross-section $\Theta(s)$ of the PVS $\Omega^i$ at $s$ and $P(s)$ is the perimeter of $\Theta(s)$. Moreover, for both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the (Dirac delta) term $\delta_\Gamma$ is a distribution concentrated on the outer lateral surfaces of the PVSs. More precisely, it is defined in terms of its \mer{integrated} action on any sufficiently smooth function $v : \Omega \rightarrow \R$ and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$ such that
\begin{equation}
  \mer{\int_{\Omega}} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
  \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v}, 
\end{equation}
under the assumption that $\xi$ is constant on each $\partial
\Theta_2^i$. 
%% , since:
%% \begin{multline}
%%   \int_{\Omega} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
%%   \int_{\Gamma} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
%%   \sum_i \int_{\Gamma^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = 
%%   \sum_i \int_{\Lambda^i} \int_{\partial \Theta_2^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \\
%%   \sum_i \int_{\Lambda^i} \left (  \xi (\overline{c} - \hat c ) \int_{\partial \Theta_2^i} \delta_\Gamma (v) \right ) =
%%   \sum_i \int_{\Lambda^i} P \xi (\overline{c} - \hat c ) \overline{v} =
%%   \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v} .
%% \end{multline}
%% \mer{MER: The previous was:
%% \begin{equation}
%%   \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \int_{\Lambda} P \xi ( \overline{c} - \hat c ) \overline{v} .
%% \end{equation}
%% }

The concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial and
perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$), entering
in~\eqref{eq:3d_pde}, represents the concentration in the perivascular
space averaged over each perivascular cross-section and is governed by 
\begin{equation}
  \partial_t (A  \hat c) - \partial_s( \hat D A \partial_s (\hat c) ) + \partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
  \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
\label{eq:1d_pde}
\end{equation}
where in $\Lambda_a, \Lambda_v$ refers to in each $\Lambda^i$ in each
of these networks. In~\eqref{eq:1d_pde}, $A = A(s)$ is defined as the
PVS cross-sectional area; i.e.,~for each $s \in \Lambda^i$ with
associated PVS $\Omega^i$, the area of the cross-section
$\Theta^i(s)$. Also, $\hat{D}$ is the effective diffusion coefficient,
and $\hat{u}$ is a convective velocity representing an (average) CSF
flow velocity in the axial direction of the perivascular spaces. To
complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
boundary conditions at the end nodes. To this end, define the set of
internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
that are connected to two or more edges and the set of end nodes as
$\mathcal{T}$, write $\hat{c}^i = \hat{c}|_{\Lambda^i}$, and let
$\mathcal{E}(y)$ denote the set of (one, two or three) edges in
$\{\Lambda^i\}$ sharing the node $y$. To impose continuity of the
concentrations, we set that the concentrations when viewed from each
edge must match at nodes:
\begin{equation}
  \hat{c}^i (y) = \hat{c}_y \qquad \foralls \Lambda^i \in \mathcal{E}(y),
  \quad \foralls y \in \mathcal{B} .
\end{equation}
Moreover, to ensure conservation of flux, we set that the flux going
in and out at nodes should add to zero, or more precisely, that for
each $y \in \mathcal{B}$:
\begin{equation}
  \sum_{i \in \mathcal{E}(y)}
  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c^i}(y) \right ) n_i(y)
  = 0
  \label{eq:flux_conservation}
\end{equation}
where the function $n_i : \mathcal{V} \rightarrow \{-1, 0, 1\}$
defines an orientation of each vertex relative to the
edges. Specifically, for an edge $\Lambda^i$ from $y_{\rm in}$ to $y_{\rm out}$
\begin{equation}
  n_i(y_{\rm in}) = 1, \quad
  n_i(y_{\rm{out}}) = -1, \quad
  n_i(y) = 0, \quad \forall y \in 
  \mathcal{V} \
  \,  \backslash \,  \{ y_{\rm in}, y_{\rm out} \}.  
\end{equation}

@@

\paragraph{Boundary and interface conditions}

On the interface $\Gamma_{\mathrm{pia}}$ between the CSF and brain, we prescribe the following interface condition modeling a semi--permeable interface, writing $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$ and $c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$:
\begin{subequations}
\begin{alignat}{2}
(-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{PAR}} &= - (- D \nabla c_{\rm{CSF}} + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n}_{\rm{PAR}} &&  \quad  \mathrm{on} \quad \pia \cup \Gamma_{\mathrm{LV}} ,  \\  
(-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{PAR}} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad  \pia \cup \Gamma_{\mathrm{LV}} .  
\end{alignat} 
\end{subequations}
Here, $\beta_{\mathrm{pia}}$ is a permeability constant between the
brain and the CSF. \mer{MER: Is the inclusion of $\Gamma_{\rm LV}$
  correct in the equation here? If so, text needs minor edits.}

For the perivascular network, the boundary nodes are collected into
$\partial \Lambda$. We assume no tracer outflow on all boundary nodes
and prescribe a homogeneous Neumann condition:
\begin{equation}
    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \partial \Lambda. 
\end{equation}

The above is supplemented with the following boundary conditions. 
\begin{alignat}{2}
(D \nabla (\phi c) - \bm u c ) \cdot \bm{n}_{\rm SAS} &= g_{\mathrm{influx}} &&  \quad  \mathrm{on} \quad \spinal  \\ 
% (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\spinal} &= 0 &&  \quad  \mathrm{on} \quad \spinal^{\mathrm{out}}  \\ 
(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\rm skull} & =   \beta c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm skull^{u}}  \\ 
(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\rm skull} & =   0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm skull^{l}}  \\ 
(-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{Pia}} &= - (-D \nabla  c_{\rm{CSF}} + \bm u_{\rm CSF} c_{\rm CSF} ) \cdot \bm{n}_{\rm{Pia}} &&  \quad  \mathrm{on} \quad \pia \cup \Gamma_{\mathrm{LV}}  \\ 
(-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n}_{\rm{Pia}} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \pia \cup \Gamma_{\mathrm{LV}}   .  
\end{alignat}

%When convenient, we also denote $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$
%and $c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$.

In the above, note that $\bm u \cdot \bm n  \neq 0$ on $\Gamma_{\rm LV} \cup \Gamma_{\rm skull^u}$ and $\bm u \cdot \bm n = 0$ on $\Gamma_{\rm SSAS} \cup \Gamma_{\rm skull^l} \cup \Gamma_{\rm Pia}$.

\paragraph*{Boundary conditions}
On $\Gamma_{\mathrm{skull}}$, the interface towards the dura and skull (\Cref{fig:concept}), we consider a constant uptake (or efflux) rate, represented by the boundary condition 
\begin{alignat}{2}
(- D \nabla ( c_{\mathrm{CSF}}) + \bm u_{\mathrm{CSF}} c_{\mathrm{CSF}}) \cdot \bm{n}   & = \beta c_{\mathrm{CSF}}  && \quad \mathrm{on} \quad \Gamma_{\mathrm{skull}^{\rm u}},\\ 
(-D \nabla c_{\rm CSF} + \bm u c_{\rm CSF}) \cdot \bm{n} & =  0   &&  \quad  \mathrm{on} \quad \Gamma_{\rm skull^{l}} , 
\end{alignat}
where $\beta$ represents a molecular outflow resistance. Here, we take $\beta = 10^{-4} \,\, \mathrm{mm}^2/\mathrm{s}$\cite{hornkjol2022csf}.

@@

Here we provide details on the numerical method used to solve the coupled system. We are given a mesh $\mathcal{T}_h $, consisting of elements denoted by $E$, of $\Omega= \brain \cup \sas$.  The collection of all interior facets to $\Omega_{\rm PAR}$ (resp.  to $\Omega_{\mathrm{CSF}}$) is  denoted by $\mathcal{F}_{h,b}$ (resp. $\mathcal{F}_{h,c}$).  The union of all interior facets $\mathcal{F}_i = \mathcal{F}_{h,b} \cup \mathcal{F}_{h,c}$. Facets lying on the pial interface are denoted by $\mathcal{F}_{\pia}$, on the skull by $\mathcal{F}_{\skull}$, and on the spinal canal by $\mathcal{F}_{\spinal}$. Note that $\mathcal{F}_i \cap \mathcal{F}_{\pia} = \emptyset $. The union of all the facets is denoted by $\mathcal{F}_h$. The restriction of $\mathcal{T}_h $ to $\Omega_{\mathrm{CSF}}$ is denoted by $\mathcal{E}_{h}$. 

@@

As boundary conditions, we impose the precomputed fluid pressure $p_{\rm CSF}$ on the networks root nodes:

@@

For the convection term, we use upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the references therein. Define 
\begin{equation*}
a_h^{\mathrm{up}}(c_h, v_h) =- \sum_{E\in\mathcal{T}_h } \int_E \bm{u} c_h \nabla v_h  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c_h\} \cdot \bm{n}_F [v_h]  + \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F |}{2} [c_h ] [v_h ] 
\end{equation*}
The discrete formulation for \eqref{eq:3d_pde} then reads 
\begin{multline}
\int_{\Omega}\frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v_h + a_h(c_h^n, v_h) + a_h^{\mathrm{up}}(c_h^n,v_h) + \int_{\Gamma_{\mathrm{pia}} } \beta_{\mathrm{pia}}[c_h^n][v_h]  \\ + \int_{\Gamma_{\rm skull^u}} \beta c_h^n v_h +  \int_{\Lambda} \xi  P (\overline{c_h^n} - \hat c_h^n) \overline{v_h}  = \int_{\Omega} f^n v_h + \int_{\spinal} \gin v_h , \quad \forall v_h \in V_h^k.
\end{multline}
For approximating the solution of the 1D equations in the network, we use the space 
\[ \hat V_h = \{ v_h \in C^0(\Lambda), 
\; v_h \vert_{\Lambda^i} \in V_h(\Lambda^i)
\}, 
\]
where $V_h(\Lambda^i)$ is the space of continuous polynomials of degree $1$ on each $\Lambda^i$,

>>>>

Also in~\eqref{eq:3d_pde}, the notation $\overline{c}$ denotes a lateral average of the concentration over the outer perivascular surfaces, defined for each centerline $\Lambda^i$ and each point $s \in \Lambda^i$ by
\begin{equation}
  \overline{c}(s) = \frac{1}{P(s)} \int_{\partial \Theta_2^i(s)} c 
\end{equation}
where $\partial \Theta_2^i(s)$ is the outer boundary of the cross-section $\Theta(s)$ of the PVS $\Omega^i$ at $s$ and $P(s)$ is the perimeter of $\Theta(s)$. Moreover, for both the periarterial and perivenous networks ($\Lambda_a, \Lambda_v$ with outer PVS boundaries $\Gamma_a, \Gamma_v$), the term $\delta_\Gamma$ is concentrated on the outer lateral surfaces of the PVSs, and defined in terms of its action on any sufficiently smooth function $v : \Omega \rightarrow \R$
%and for each $\Lambda^i \subseteq \Lambda$ with PVS $\Omega^i$ and outer lateral surface $\Gamma^i = \partial \Omega^i_2$
such that
\begin{equation}
  \langle \xi (\overline{c} - \hat c ) \delta_\Gamma, v \rangle =
  \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v}, 
\end{equation}
under the assumption that $\xi$ is constant on each $\partial
\Theta_2^i$, see also~\cite{masri2024discontinuous}.
%% , since:
%% \begin{multline}
%%   \int_{\Omega} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
%%   \int_{\Gamma} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) =
%%   \sum_i \int_{\Gamma^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = 
%%   \sum_i \int_{\Lambda^i} \int_{\partial \Theta_2^i} \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \\
%%   \sum_i \int_{\Lambda^i} \left (  \xi (\overline{c} - \hat c ) \int_{\partial \Theta_2^i} \delta_\Gamma (v) \right ) =
%%   \sum_i \int_{\Lambda^i} P \xi (\overline{c} - \hat c ) \overline{v} =
%%   \int_{\Lambda} P \xi (\overline{c} - \hat c ) \overline{v} .
%% \end{multline}
%% \mer{MER: The previous was:
%% \begin{equation}
%%   \xi (\overline{c} - \hat c ) \delta_\Gamma (v) = \int_{\Lambda} P \xi ( \overline{c} - \hat c ) \overline{v} .
%% \end{equation}
%% }

The concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial and
perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$), entering
in~\eqref{eq:3d_pde}, represents the concentration in the perivascular
space averaged over each perivascular cross-section and is governed by 
\begin{equation}
  \partial_t (A  \hat c) - \partial_s( \hat D A \partial_s (\hat c) ) + \partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
  \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
\label{eq:1d_pde}
\end{equation}
where in $\Lambda_a, \Lambda_v$ refers to in each $\Lambda^i$ in each
of these networks. In~\eqref{eq:1d_pde}, $A = A(s)$ is defined as the
PVS cross-sectional area; i.e.,~for each $s \in \Lambda^i$ with
associated PVS $\Omega^i$, the area of the cross-section
$\Theta^i(s)$. Also, $\hat{D}$ is the effective diffusion coefficient,
and $\hat{u}$ is a convective velocity representing an (average) CSF
flow velocity in the axial direction of the perivascular spaces. To
complete~\eqref{eq:1d_pde}, we prescribe bifurcation conditions at
internal nodes of the perivascular networks $\Lambda_a, \Lambda_v$ and
boundary conditions at the end nodes. To this end, define the set of
internal nodes $\mathcal{B} \subset \mathcal{V}$ as the set of nodes
that are connected to two or more edges and the set of end nodes as
$\mathcal{T}$, write $\hat{c}^i = \hat{c}|_{\Lambda^i}$, and let
$\mathcal{E}(y)$ denote the set of (one, two or three) edges in
$\{\Lambda^i\}$ sharing the node $y$. To impose continuity of the
concentrations, we set that the concentrations when viewed from each
edge must match at nodes:
\begin{equation}
  \hat{c}^i (y) = \hat{c}_y \qquad \foralls \Lambda^i \in \mathcal{E}(y),
  \quad \foralls y \in \mathcal{B} .
\end{equation}
Moreover, to ensure conservation of flux, we set that the flux going
in and out at nodes should add to zero, or more precisely, that for
each $y \in \mathcal{B}$:
\begin{equation}
  \sum_{i \in \mathcal{E}(y)}
  \left (\hat{D}^i A^i \partial_s \hat{c}^i (y) - A^i \hat u^i(y) \hat{c^i}(y) \right ) n_i(y)
  = 0
  \label{eq:flux_conservation}
\end{equation}
where the function $n_i : \mathcal{V} \rightarrow \{-1, 0, 1\}$
defines an orientation of each vertex relative to the
edges. Specifically, for an edge $\Lambda^i$ from $y_{\rm in}$ to $y_{\rm out}$
\begin{equation}
  n_i(y_{\rm in}) = 1, \quad
  n_i(y_{\rm{out}}) = -1, \quad
  n_i(y) = 0 \quad \foralls y \in 
  \mathcal{V} \backslash \{ y_{\rm in}, y_{\rm out} \} .  
\end{equation}

\paragraph{Interface conditions between the brain and CSF spaces}

On the interface between the brain and the CSF spaces
$\Gamma_{\mathrm{pia}} \cup \Gamma_{\rm LV}$, we prescribe the
following interface conditions, which represent a semi--permeable
interface, writing $c|_{\Omega_{\rm PAR}} = c_{\rm PAR}$ and
$c|_{\Omega_{\rm CSF}} = c_{\rm CSF}$:
%% \begin{subequations}
%% \begin{alignat}{2}
%% (-D \nabla (\phi c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} &= - (- D (\phi \nabla c_{\rm{CSF}}) + \bm u_{\rm{CSF}} c_{\rm{CSF}} ) \cdot \bm{n} &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\  
%% (-D \nabla ( c_{\rm{PAR}}) + \bm u_{\rm{PAR}} c_{\rm{PAR} }) \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  
%% \end{alignat} 
%% \end{subequations}
\begin{subequations}
  \begin{alignat}{2}
  (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n}
  &= - (-D \nabla (\phi c) + \bm u c)|_{\Omega_{\rm CSF}} \cdot \bm{n}
  &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  \\  
  (- D \nabla (\phi  c) + \bm u c)|_{\Omega_{\rm PAR}} \cdot \bm{n} & = \beta_{\rm{pia}} (c_{\rm PAR} - c_{\mathrm{CSF}}) &&  \quad  \mathrm{on} \quad \Gamma_{\rm pia} \cup \Gamma_{\mathrm{LV}} ,  
\end{alignat} 
  \label{eq:interface:3d}%
\end{subequations}%
where $\bm{n} = \bm{n}_{\rm PAR}$ is the normal vector field defined
over the interface, outward-pointing when viewed from $\Omega_{\rm
  PAR}$, and $\beta_{\mathrm{pia}} \geqslant 0$ is a permeability
constant.

\paragraph{Boundary conditions: outer SAS interfaces and perivascular ends}

We supplement~\eqref{eq:3d_pde} and~\eqref{eq:interface:3d} with the following boundary conditions representing a given molecular influx at $\Gamma_{\rm SSAS}$, molecular efflux at a constant update rate $\beta_{\rm exit}$ at $\Gamma_{\rm AM-U}$, and no influx or efflux elsewhere from the CSF spaces.
\begin{alignat}{2}
(D \nabla (\phi c) - \bm u c ) \cdot \bm{n} &= g_{\mathrm{influx}} &&  \quad  \mathrm{on} \quad \Gamma_{\rm SSAS},  \\ 
% (-D \nabla (\phi c) + \bm u c ) \cdot \bm{n}_{\Gamma_{\rm SAS}} &= 0 &&  \quad  \mathrm{on} \quad \Gamma_{\rm SAS}^{\mathrm{out}}  \\ 
(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = \beta c  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-U},  \\ 
(-D \nabla (\phi c) + \bm u c ) \cdot \bm{n} & = 0  &&  \quad  \mathrm{on} \quad \Gamma_{\rm AM-L} \discuss{\cup \Gamma_{\rm SC}}  .
\end{alignat} 
where again $\bm n$ denotes the outward-pointing boundary normal.

%In the above, note that $\bm u \cdot \bm n  \neq 0$ on $\Gamma_{\rm LV} \cup \Gamma_{\rm skull^u}$ and $\bm u \cdot \bm n = 0$ on $\Gamma_{\rm SSAS} \cup \Gamma_{\rm skull^l} \cup \Gamma_{\rm Pia}$.

For the perivascular networks, we assume no molecular efflux at all
end nodes $\mathcal{T}$ via a homogeneous Neumann condition
augmenting~\eqref{eq:1d_pde}:
\begin{equation}
    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \mathcal{T}.
\end{equation}

@@

@@

Here we provide details on the numerical method used to solve the coupled system. We are given a mesh $\mathcal{T}_h $, consisting of elements denoted by $E$, of $\Omega= \Omega_{\rm PAR} \cup \Omega_{\rm CSF}$.  The collection of all interior facets to $\Omega_{\rm PAR}$ (resp.  to $\Omega_{\mathrm{CSF}}$) is  denoted by $\mathcal{F}_{h,b}$ (resp. $\mathcal{F}_{h,c}$).  The union of all interior facets $\mathcal{F}_i = \mathcal{F}_{h,b} \cup \mathcal{F}_{h,c}$. Facets lying on the pial interface are denoted by $\mathcal{F}_{\Gamma_{\rm pia}}$, on the skull by $\mathcal{F}_{\Gamma_{\rm skull}}$, and on the spinal canal by $\mathcal{F}_{\Gamma_{\rm SAS}}$. Note that $\mathcal{F}_i \cap \mathcal{F}_{\Gamma_{\rm pia}} = \emptyset $. The union of all the facets is denoted by $\mathcal{F}_h$. The restriction of $\mathcal{T}_h $ to $\Omega_{\mathrm{CSF}}$ is denoted by $\mathcal{E}_{h}$. 

@@

As boundary conditions, we impose the precomputed fluid pressure $p_{\rm CSF}$ on the networks root nodes:
\begin{equation}
    p_{\rm PVS} = p_{\rm CSF} \quad \text{ on } \Lambda^{\rm out}
\end{equation}

@@

For the convection term, we use upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the references therein. Define 
\begin{equation*}
a_h^{\mathrm{up}}(c_h, v_h) =- \sum_{E\in\mathcal{T}_h } \int_E \bm{u} c_h \nabla v_h  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c_h\} \cdot \bm{n}_F [v_h]  + \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F |}{2} [c_h ] [v_h ] 
\end{equation*}
The discrete formulation for \eqref{eq:3d_pde} then reads 
\begin{multline}
\int_{\Omega}\frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v_h + a_h(c_h^n, v_h) + a_h^{\mathrm{up}}(c_h^n,v_h) + \int_{\Gamma_{\mathrm{pia}} } \beta_{\mathrm{pia}}[c_h^n][v_h]  \\ + \int_{\Gamma_{\rm skull^u}} \beta c_h^n v_h +  \int_{\Lambda} \xi  P (\overline{c_h^n} - \hat c_h^n) \overline{v_h}  = \int_{\Omega} f^n v_h + \int_{\Gamma_{\rm SAS}} \gin v_h , \quad \forall v_h \in V_h^k.
\end{multline}
For approximating the solution of the 1D equations in the network, we use the space 
\[ \hat V_h = \{ v_h \in C^0(\Lambda), 
\; v_h \vert_{\Lambda^i} \in V_h(\Lambda^i)
\}, 
\]
where $V_h(\Lambda^i)$ is the space of continuous polynomials of degree $1$ on each $\Lambda^i$,
